---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---
## Loading and preprocessing the data
```{r eval=T,echo=T,warning=F,message=F}
# Loading libraries
library(dplyr)
library(ggplot2)
library(lubridate)
library(kableExtra)
```

## Downloading the data if the zip file does not exist, extracting and reading the csv file.

```{r eval=T,echo=T,warning=F,message=F}
setwd("~\\GitHub\\RepData_PeerAssessment1")
if(!exists("data.zip")){
download.file("https://d396qusza40orc.cloudfront.net/repdata%2Fdata%2Factivity.zip","data.zip")
unzip("data.zip")
}
activity<-read.csv("activity.csv",header = T) %>% as_tibble()
# Converting date to correct format
activity$date <-ymd(activity$date)
```
## Imputing the missing values
```{r eval=T,echo=T,warning=F,message=F}
nrow(activity)/nrow(activity %>% filter(is.na(activity$steps)))
### There are about 7% missing values. We will fill those with median values
activity$steps[which(is.na(activity$steps))]<-median(activity$steps,na.rm = T)
```

## Total number of steps per day is calculated

```{r eval=T,echo=T,warning=F,message=F}
daily_activity<-activity %>% group_by(date) %>% summarise(steps=sum(steps))
head(daily_activity)

boxplot(daily_activity$steps,xlab = "Total number of steps taken per day")

```

## Histogram of the total number of steps taken each day

```{r eval=T,echo=T,warning=F,message=F}

hist(daily_activity$steps,xlab = "Total number of steps taken per day")
```

## Mean and median number of steps taken each day

```{r eval=T,echo=T,warning=F,message=F}
"Mean of daily total steps is " %>% paste( mean(daily_activity$steps,na.rm = T))
"Median of daily total steps is " %>% paste( median(daily_activity$steps,na.rm = T))
```

## Time series plot of the average number of steps taken

```{r eval=T,echo=T,warning=F,message=F}
plot(x=daily_activity$date,y=daily_activity$steps,type="l")
```

## The 5-minute interval that contains the maximum number of steps

```{r eval=T,echo=T,warning=F,message=F}
activity %>% group_by(interval) %>% summarise(steps=mean(steps)) %>% top_n(1,wt=steps)
```

## Are there differences in activity patterns between weekdays and weekends?

```{r eval=T,echo=T,warning=F,message=F}
# Marking weekdays and weekends
daily_activity<-daily_activity %>% mutate(wd=ifelse(weekdays(date)=="Sunday"|weekdays(date)=="Saturday","Weekends","Weekdays"))


plot(daily_activity$wd %>% as.factor(),daily_activity$steps)
abline(h=mean(daily_activity$steps),col="red")


```

## Panel plot comparing the average number of steps taken per 5-minute interval across weekdays and weekends

```{r eval=T,echo=T,warning=F,message=F}
temp<-activity %>% mutate(wd=ifelse(weekdays(date)=="Sunday"|weekdays(date)=="Saturday","Weekends","Weekdays")) %>% group_by(interval,wd,date) %>% summarise(steps=mean(steps))

temp.wd<-temp %>% filter(wd=="Weekdays")
temp.we<-temp %>% filter(wd=="Weekends")

par(mfrow=c(2,2))
plot(x=temp.wd$date,y=temp.wd$steps,type="p")
plot(x=temp.we$date,y=temp.we$steps,type="p")

```
